Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-06-15 snapshot_system/snapshot.tex
modify the analysi

\section{Snapshot Deduplication}
\label{sect:dedupe}
%snapshot representation

@@

Our  deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud.
This process performs   the duplication in two categories: \textit{Inner-VM} and \textit{Cross-VM}. 
We discuss the integration of several strategies as follows and analyze its effectiveness.

\subsection{Inner and corss VM deduplication}

Inner-VM duplication can be effecive within  VM's snapshots because the majority of data are unchanged during each backup period. 
Localizing the snapshot data deduplication witin  a VM first
increases data independency between different VM backups,
simplifies snapshot management and statistics collection,
and facilitates parallel execution of snapshot operations.
On the other hand, cross-VM duplication can also be desirable mainly due to widely-used software and libraries. 
As the result, different VMs tend to backup large amount of highly similar data.
We integrate those strategies together as follows.
%Our multi-level pipeline process can minimize 
%the cost of deduplication while maximize the its efficiency at each level,
%and it is parallel since each segment is processed independently.

@@

\subsection{Popular Chunk Management and VM-oriented Fault Isolation}
Our objective for fault isolation is to minimize the number of VMs affected when there are failures
in the cluster.  The inner-VM deduplication does not depend on any global service and the comparison
for each VM is localized within the parent and the current snapshot.
Thus there is no data dependence between VMs.
For cross-VM deduplication, there is a data dependence among VMs and we would minimize the faiulre impact
of shared blocks by adding extra replcas of those shared blocks.

This section analyzes the choice of popular blocks and its impact on the deduplication efficiency.
It also  compares the  fault resilience of our VM-centric deduplication approach with a standand approach using 
global deduplication.

\subsubsection{Impact of CDS deduplication}

Our empirical study based on VM images from production environment\cite{ieeecloud} showed that the
frequency of data duplication follows Zipf-like distribution\cite{zipf},
with the exponent $\alpha$ between 0.65 ~ 0.70.
As a result, it can be proved that deduplication efficiency of CDS index is scalable:

For the Zipf-like distribution, an approximation to the sum of the first $n$ 
elements of the distribution can be derived as follows:
\begin{equation}
\sum_{i=1}^{n}\frac{1}{i^\alpha}\approx \int_{1}^{n}\frac{1}{x^\alpha}\mathrm{d}x=\frac{x^{1-\alpha}}{1-\alpha}=\frac{n^{1-\alpha}}{1-\alpha}\;  for\;  \alpha<1
\end{equation}

So the cumulative distribution function for a CDS holding top $S_c$ fingerprints
of global index with size $S_g$ is:

\begin{equation}
  E = (S_c / S_g)^{1-\alpha} \;  for\;  \alpha<1
\end{equation}

%[Describe the dedup efficiency model in detail]
Let $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by CDS, $D$ be the amount of data on each node, and $B$ be the average block size. Then $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}

By replacing $S_c$ and $S_g$ in the first formula, the deduplication efficiency becomes:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}

Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of CDS is only controlled by the its memory usage.

>>>>

\section{VM-centric Snapshot Deduplication}
\label{sect:dedupe}
%snapshot representation

@@

A deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud.
The tradition approach compares all fingerprints and 
reserve one copy for all of its deduplicates.
We call this as  the VM-oblivious (VO) approach.
In comparison, we analyze the design of
a VM-centric approach (VC) which differenates duplicates within a VM and cross VMs,
and conducts \textit{Inner-VM} and \textit{cross-VM} detection.

\subsection{Inner and cross VM deduplication}

Inner-VM duplication can be effecive within  VM's snapshots there
are a large amount of duplicates existing among the snapshots within
each VM.  Localizing the snapshot data deduplication witin  a VM first
increases data independency between different VM backups,
simplifies snapshot management and statistics collection,
and facilitates parallel execution of snapshot operations.
On the other hand, cross-VM duplication can also be desirable mainly due to widely-used software and libraries. 
As the result, different VMs tend to backup large amount of highly similar data.
We integrate those strategies together as follows.
%Our multi-level pipeline process can minimize 
%the cost of deduplication while maximize the its efficiency at each level,
%and it is parallel since each segment is processed independently.

@@

\subsection{Popular Chunk Management}
%VM-oriented Fault Isolation}
%Our objective for fault isolation is to minimize the number of VMs affected when there are failures
%in the cluster.  
%The inner-VM deduplication does not depend on any global service and the comparison
%for each VM is localized within the parent and the current snapshot.
%Lack of data dependence between VMs brings 
With cross-VM deduplication, shared data blocks
create an artifical dependence  among VMs and failure of one shared block
affects many VMs. Thus we only maintain the index for a small set of popular chunks.
The mnagement for popular data chunks contains two aspects.
\begin{itemize}
\item Compute and select top-$k$ most popular chunks.
The popularity of a chunk is the number  of data blocks from different VMs
that are duplicates of this block after the inner VM deduplication.
This number can be computed perioidically in a weekly basis.
Once the popularity of all data blocks is collected, the system only maintains the top $k$
most popular blocks.
\item  Top $k$ number is chosen to be small while 
our analysis belows will show that the algorithm can still deliver competitive deduplication effiency.
Since $k$ is small and these top $k$ blocks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular blocks to enhanace the fault resilence.
We will provide an alaysis in the next subsection.
\end{itemize}

This section analyzes how value   $k$ impacts on the deduplication efficiency.
%It also  compares the  fault resilience of our VM-centric deduplication approach with a standand approach using 
%global deduplication.
%\subsubsection{Impact of CDS deduplication}
The analysis is based on the our study 
using Alibaba's production user data~\cite{ieeecloud} 
which shows that the popularity of data blocks after inner VM deduplication follows 
a Zipf-like distribution\cite{zipf} and its
exponent $\alpha$ is ranged between 0.65  and  0.70.
%As a result, it can be proved that deduplication efficiency of CDS index is scalable:

\begin{table}[ht]
\centering
\begin{tabular}{|p{1.25cm}|p{6.5cm}|}
\hline
$b$ &  the total amount of data blocks\\ 
\hline
$b_u$ &  the total amount of unique fingerprints after inner VM  deduplication\\
\hline
$C_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$s$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$s$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$p$ & the number of nodes in the cluster\\
\hline
$D$ & the amount of unique data on each node\\
\hline
$B$ & the average block size\\
\hline
$m$ & memory size on each node used by VC\\ 
\hline
$F$ & the size of an polular data index entry\\
\hline
\end{tabular}
\caption{Modeling  symbols.}
\label{tab:symbol}
\end{table}

Let $b$ be the total amount of data blocks after localized deduplication, 
$b_u$ be the total amount of fingerprints 
in the global index after complete deduplication, 
$C_i$ be the frequency for the $i$th most popular fingerprint. 
By Zipf-like distribution, $C_i = \frac{C_1}{i^\alpha}.$
Since $ \sum_{i=1}^{b_u}C_i = b$,
\[
C_1 \sum_{i=1}^{b_u}\frac{1}{i^\alpha} = b
\]
Given $\alpha <1$, $C_1$ can be approximated with integration:
\begin{equation}
C_1=\frac{b(1-\alpha)}{b_u^{1-\alpha}}.
\end{equation}

%We define deduplication efficiency $e$ to be the proportion of raw data that can be deduplicated,
%In complete deduplication scenario, it deduplication efficiency $e_c$ is:
%\begin{equation}
%e_c = \frac{b-b_u}{b} = 1 - \frac{b_u}{b}
%\end{equation}

The  $k$ most popular fingerprints can cover the following number of blocks.
\[
C_1 \sum_{i=1}^{k}\frac{1}{i^\alpha} \approx  
C_1 \int_{1}^{k}\frac{1}{x^\alpha} dx  \approx  C_1\frac{  k^{1-\alpha}} {1-\alpha}.
\]

Deduplication efficiency of VC using top $k$ popular blocks
is the percentage of duplicates that can be detected:  
\begin{equation}
\begin{split}
e_k &= 
\frac{ b(1-s) + C_1\frac{  k^{1-\alpha}} {1-\alpha}} 
{b(1-s)  + sb - b_u }\\
&= 
\frac{ (1-s) + s (\frac{k}{b_u})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\end{split}
\end{equation}

%\begin{equation}
%\begin{split}
%e_k &= (\sum_{i=1}^{k}C_i - k)/{b} = (b * \frac{1-\alpha}{b_u^{1-\alpha}} * \sum_{i=1}^{k}\frac{1}{i^\alpha} - k)/b \\
%&\approx \frac{b*\frac{1-\alpha}{b_u^{1-\alpha}}*\frac{k^{1-alpha}}{1-\alpha} - k}{b}
%&\approx (\frac{k}{b_u})^{1-\alpha} - \frac{k}{b} \approx (\frac{k}{b_u})^{1-\alpha},\; (since\; k \ll b)
%\end{split}
%\end{equation}

% For the Zipf-like distribution, an approximation to the sum of the first $n$ 
% elements of the distribution can be derived as follows:
% \begin{equation}
% \sum_{i=1}^{n}\frac{1}{i^\alpha}\approx \int_{1}^{n}\frac{1}{x^\alpha}\mathrm{d}x=\frac{x^{1-\alpha}}{1-\alpha}=\frac{n^{1-\alpha}}{1-\alpha}\;  for\;  \alpha<1
% \end{equation}

% So the cumulative distribution function for a CDS holding top $S_c$ fingerprints
% of global index with size $S_g$ is:

% \begin{equation}
%   E = (S_c / S_g)^{1-\alpha} \;  for\;  \alpha<1
% \end{equation}

%[Describe the dedup efficiency model in detail]
Let $p$ be the number of physical machines in the cluster, $m$ be the memory on each node used by the popular
index, $F$ be the size of an index entry,
$D$ be the amount of unique data on each phyiscal machine, and 
$B$ be the average block size. We store the popular index using a distribued shared memory hashtable
such as MemCacheD.  Then $k$ and $b_u$ can be expressed as:
$
k = p*m/F$, and $b_u = p*D/B.
$

The overall deduplication effiency is
\[
\frac{ (1-s) + s (\frac{m*B}{D*F})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\]
where $(\frac{m*B}{D*F})^{1-\alpha}$ represents the percentage of the remaining blocks detected as duplicates
after inner VM deduplication. 
%Then the deduplication efficiency of CDS approach using top k most popular fingerprints is:
%\begin{equation}
%  e_k = (\frac{m*B}{F*D})^{1-\alpha}
%\end{equation}

When the number of machines at each cluster increases, the number of total VMs increases.
Then $k$ increases since more memory is available to host the popular block index.
But for each physical machine, the number of VMs remains the same, and thus
$D$ is  a constant. Then  the overall deduplication efficiency of VC remains
a constant.

% is mainly affected by the its memory usage,
%and the Zipf-like distribution parameter $\alpha$.
%From the memory usage perspective, the more memory we give to CDS index, the more duplicates it can found.
%From the data characteristics perspective, if the system tends to backup arbitrarily random data, 
%$\alpha$ would be small and thus CDS approach would not be effective. But in a VM cluster all VMs
%have a large portion of similar data, thus leading to a bigger $\alpha$. The more VM join the cluster,
%the better deduplication efficiency our CDS deduplication approach will have.

\begin{table}
    \begin{tabular}{llll}
    Data size (GB) & 1\%    & 2\%    & 4\%    \\
    14.6           & 18.6\% & 22.1\% & 31.4\% \\
    28.1           & 19.5\% & 26.2\% & 38.8\% \\
    44.2           & 21.7\% & 26.5\% & 36\%   \\
    61.6           & 23.2\% & 32.9\% & 35\%   \\
    74.2           & 23.6\% & 33.6\% & 37.5\% \\
    \end{tabular}
    \caption{Deduplication effectiveness of top k\% of global index}
    \label{tab:cds}
\end{table}
